---
title: "plots_maps"
author: "Renjie Wei"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(plotly)
library(skimr)
library(rpart)
library(corrplot)
library(tibbletime)
library(lubridate)
library(smooth)
library(grid)
library(gridExtra)
library(magrittr)
library(cowplot)
library(ggthemes)


library(broom)
library(highcharter)
library(rworldmap)
library(gganimate)
library(transformr)
library(corrplot)
library(leaps)
library(kableExtra)

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = F
)

conv_color = "#F17C67"
org_color = "#7BA23F"

```


From Renjie, trying to do more fancy plots and interactive maps.


```{r }
# from Beibei
avo_df = 
  read_csv("data/avocado.csv") %>% 
  janitor::clean_names() %>% 
  select(-1) %>% 
  separate(date, c("year", "month", "day"), remove = FALSE) %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)
  ) %>% 
  rename(
    small = x4046,
    large = x4225,
    extra_large = x4770,
  ) 

avo_tidy = 
  avo_df %>% 
  pivot_longer(
    small:extra_large,
    names_to = "fruit_size",
    values_to = "quantity_sold"
  ) %>% 
  pivot_longer(
    total_bags:x_large_bags,
    names_pattern = "(.*)_bags",
    names_to = "bag_type",
    values_to = "bag_sold",
  ) %>% 
  mutate(
    bag_type = recode(bag_type, x_large = "extra_large")
  )


gdp_df = 
  read_csv("data/gdp-by-state.csv")
```


# Types of Avocados:

In this section we will analyze the different types of avocados that we have in this dataset. Basically, we have two types of avocados:

  + Conventional
  + Organic
  
**Summary:**

  + Organic avocados: Based on the price changes throughout time we can see that they are more expensive.
  + Conventional avocados: Based on price changes throughout time we can see that they are less expensive.

## Trend in Price:

```{r price_vs_type}
avo_df %>% 
  ggplot(aes(x=average_price, fill=type)) + 
  geom_density() + 
  facet_wrap(~type) + 
  theme_minimal() + 
  theme(plot.title=element_text(hjust=0.5), legend.position="bottom") + 
  labs(title="Avocado Price by Type") 

```


```{r price_trend_by_type}
# Sort the dates
avo_df <- avo_df[order(as.Date(avo_df$date, format="%Y-%m-%d")),]


price_trend <- 
  avo_df %>% 
  select(date, average_price, type) %>%
  ggplot(aes(x=date, y=average_price)) +
  geom_area(aes(color=type, fill=type), alpha = 0.3, position = position_dodge(0.8)) + 
  theme_minimal()

price_trend


ggplot(data = avo_df, aes(x = date, y = average_price, col=type)) +
  geom_line() +
  facet_wrap(~ type) + 
  theme_minimal() + theme(legend.position="bottom") +
  theme_minimal()
```


## Relationship between Prices and Total Volume:

贝贝做过了说辣鸡的是这个么...

```{r volume_price}
# filter dataset

organic <- avo_df %>% select(date, average_price, type, total_volume) %>% filter(type == "organic")
conventional <- avo_df %>% select(date, average_price, type, total_volume) %>% filter(type == "conventional")


organic <- as_tbl_time(organic, index=date)
organic <- as_period(organic, '1 month')

conventional <- as_tbl_time(conventional, index=date)
conventional <- as_period(conventional, '1 month')
```

```{r price_volume_plot}
conventional_monthly <- 
  conventional %>%
ggplot(aes(x=date, y=average_price)) + 
  geom_line(color=conv_color) + 
  theme(plot.title=element_text(hjust=0.5)) + 
  labs(title="Conventional Avocados") + geom_hline(yintercept=max(conventional$average_price), linetype="dashed", color = "red") + 
  geom_hline(yintercept=min(conventional$average_price), linetype="dashed", color = "blue") +
  theme_minimal()


conventional_volume <- 
  conventional %>%
  ggplot(aes(x=date, y=total_volume)) + 
  geom_bar(stat='identity', fill=conv_color, color=conv_color) + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_smooth(method="loess", color="red") +
  theme_minimal()

organic_monthly <- 
  organic %>% 
  ggplot(aes(x=date, y=average_price)) + 
  geom_line(color=org_color) + 
  theme(plot.title=element_text(hjust=0.5)) + 
  labs(title="Organic Avocados") + geom_hline(yintercept=max(organic$average_price), linetype="dashed", color = "red") + 
  geom_hline(yintercept=min(organic$average_price), linetype="dashed", color = "blue") +
  theme_minimal()

organic_volume <- 
  organic %>%
  ggplot(aes(x=date, y=total_volume)) + 
  geom_bar(stat='identity', fill=org_color ,color=org_color) + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_smooth(method="loess", color="red") +
  theme_minimal()

plot_grid(conventional_monthly, organic_monthly,conventional_volume, organic_volume, nrow=2, ncol=2)
```


 + Volume: Each volume **peak** is a signal for an upcoming **drop** in avocado prices.

# Seasonal Patterns?:

If there any repeating trends in which the avocado price tend to increase?

## Price by year


```{r sesonal_df}

seasonal_df =  
  avo_df %>% 
  mutate(
    month_year <- format(as.Date(date), "%Y-%m"),
    month <- format(as.Date(date), "%m"),
    format(as.Date(date), "%Y"),
    monthabb = sapply(month, function(x) month.abb[as.numeric(x)]),
    monthabb = factor(monthabb, levels = month.abb)
  )



# sesonal plot
seasonal_df %>% 
  ggplot(aes(x = average_price, fill = as.factor(year))) + 
  geom_density(alpha = .5) + 
  theme_minimal() +
  facet_wrap(~ year) + theme(plot.title=element_text(hjust=0.5)) + 
  guides(fill = FALSE) + labs(title="Distribution of Prices by year", x = 'Average Price', y = 'Density')

```

## Price trend by type 

```{r price_trend_total}

conv_patterns = 
  seasonal_df %>% 
  select(monthabb, average_price, type) %>% 
  filter(type == "conventional") %>%
  group_by(monthabb) %>% 
  summarize(avg=mean(average_price)) %>%
  ggplot(aes(x=monthabb, y=avg)) + 
  geom_point(color="#9E7A7A", aes(size=avg)) + 
  geom_line(group=1, color= conv_color) + 
  theme_minimal() + 
  theme(legend.position="none", plot.title=element_text(hjust=0.5)) + 
  labs(title="Conventional Avocados", x="Month", y="Average Price")


org_patterns = 
  seasonal_df %>% 
  select(monthabb, average_price, type) %>% 
  filter(type == "organic") %>%
  group_by(monthabb) %>% 
  summarize(avg=mean(average_price)) %>%
  ggplot(aes(x=monthabb, y=avg)) + 
  geom_point(color="#9E7A7A", aes(size=avg)) + 
  geom_line(group=1, color=org_color) + 
  theme_minimal() + 
  theme(legend.position="none", plot.title=element_text(hjust=0.5)) + 
  labs(title="organic Avocados", x="Month", y="Average Price")

plot_grid(conv_patterns, org_patterns, nrow=2)
```

And to look closer, we create the seasonality patterns in each year

```{r price_trend_seasonal}
conv_pat_yearly =
  seasonal_df %>% 
  select(year, monthabb, average_price, type) %>% 
  filter(type == "conventional", year == c("2015", "2016", "2017")) %>%
  group_by(year, monthabb) %>% 
  summarize(avg=mean(average_price)) %>%
  ggplot(aes(x=monthabb, y=avg)) + 
  geom_point(color="#9E7A7A") + 
  geom_line(group=1, color=conv_color) + 
  facet_wrap(~as.factor(year)) + 
  theme_minimal() + 
  theme(plot.title=element_text(hjust=0.5), axis.text.x = element_text(angle = 90)) + 
  labs(title="Seasonal Fluctuations \n Convenctional Avocados", x="Month", y="Average Price")

org_pat_yearly =
  seasonal_df %>% 
  select(year, monthabb, average_price, type) %>% 
  filter(type == "organic", year == c("2015", "2016", "2017")) %>%
  group_by(year, monthabb) %>% 
  summarize(avg=mean(average_price)) %>%
  ggplot(aes(x=monthabb, y=avg)) + 
  geom_point(color="#9E7A7A") + 
  geom_line(group=1, color=org_color) + 
  facet_wrap(~as.factor(year)) + 
  theme_minimal() + 
  theme(plot.title=element_text(hjust=0.5), axis.text.x = element_text(angle = 90)) + 
  labs(title="Seasonal Fluctuations \n Convenctional Avocados", x="Month", y="Average Price")
 

plot_grid(conv_pat_yearly, org_pat_yearly, nrow=2)
```

Make some barplots to show the changes between year.


```{r diff_bar}
r_avg = 
  seasonal_df %>% 
  group_by(year, monthabb) %>%  
  select(type, year, monthabb, average_price) %>% 
  filter(type == "conventional", year == c("2015", "2016", "2017")) %>%
  group_by(year, monthabb) %>%
  summarize(avg=mean(average_price))



wider_df =
  r_avg %>% 
  pivot_wider(
    names_from = year,
    values_from = avg
  )


colnames(wider_df) = c("Months", "First_year", "Second_year", "Third_year")

wider_df = 
  wider_df %>% 
  mutate(
    first_pct = (Second_year - First_year)/ First_year,
    second_pct = (Third_year - Second_year)/ Second_year,
    first_sign = ifelse(first_pct > 0, "Positive", "Negative"),
    second_sign = ifelse(second_pct > 0, "Positive", "Negative")
  )


pct_dif_1 =
  wider_df %>% 
  select(Months, first_pct, first_sign) %>%
  ggplot(aes(fill=first_sign)) + 
  geom_bar(stat='identity', aes(x = Months, y = round(first_pct,4) * 100)) + 
  theme_minimal() + 
  theme(axis.text.x=element_text(angle=90), legend.position="bottom") + 
  labs(x="Month", y="Difference(%)") + 
  guides(fill=guide_legend(title="Diff Status")) + 
  scale_fill_manual(values=c("#EB7A77", "#91AD70")) +
  labs(title="Conventional Avocado Price changes \n (2015 - 2016)", x="Months", y="Price")

pct_dif_2 =
  wider_df %>% 
  select(Months, second_pct, second_sign) %>%
  ggplot(aes(fill=second_sign)) + 
  geom_bar(stat='identity', aes(x = Months, y = round(second_pct,4) * 100)) + 
  theme_minimal() + 
  theme(axis.text.x=element_text(angle=90), legend.position="bottom") + 
  labs(x="Month", y="Difference(%)") + 
  guides(fill=guide_legend(title="Diff Status")) + 
  scale_fill_manual(values=c("#EB7A77", "#91AD70")) +
  labs(title="Conventional Avocado Price changes \n (2016 - 2017)", x="Months", y="Price")



plot_grid(pct_dif_1, pct_dif_2, ncol=2)
```

# Regional Analysis

```{r}

```

