---
title: "EDA"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(plotly)

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Import avocado data. 
```{r}
avo_df = 
  read_csv("data/avocado.csv") %>% 
  janitor::clean_names() %>% 
  select(-1) %>% 
  separate(date, c("year", "month", "day"), remove = FALSE) %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)
  ) %>% 
  rename(
    small = x4046,
    large = x4225,
    extra_large = x4770,
  ) 

avo_tidy = 
  avo_df %>% 
  pivot_longer(
    small:extra_large,
    names_to = "fruit_size",
    values_to = "quantity_sold"
  ) %>% 
  pivot_longer(
    total_bags:x_large_bags,
    names_pattern = "(.*)_bags",
    names_to = "bag_type",
    values_to = "bag_sold",
  ) %>% 
  mutate(
    bag_type = recode(bag_type, x_large = "extra_large")
  )

avo_tidy
```

Description: \
`year`: 2015-2018 \
`month`: 1-12 \
`day`: 1-31 \
`type`: conventional, organic \
`fruit_size`: small, large, extra_large \
`bag_type`: total, small, large, extra_large \

```{r}
gdp_df = 
  read_csv("data/gdp-by-state.csv") %>% 
  janitor::clean_names() %>% 
  select(-1, -x2013, -x2014) %>% 
  pivot_longer(
    x2015:x2017,
    names_prefix = "x",
    names_to = "year",
    values_to = "gdp"
  )

gdp_df
```

Description: \
`year`: 2015-2017 \

```{r include=FALSE}
# number of regions slightly different  
sort(unique(gdp_df$area))
sort(unique(gdp_df$area))
```

wo jue de ke yi zhao you guan xi de che yi che ?
https://www.medicalnewstoday.com/articles/270406#benefits
https://pdf.usaid.gov/pdf_docs/PA00KP28.pdf

yao bu zhe li zai gao dian data 
fao.org/faostat/en/#search/Avocados


https://quickstats.nass.usda.gov/results/8A9760E3-BDB0-3A88-B014-DA81BA0845BD


Volume consumption by year: conventional vs. organic
```{r fig.width = 5, fig.height=5}
avo_df %>% 
  group_by(year, type) %>% 
  summarise(sum_volume = sum(total_volume)) %>% 
  pivot_wider(
    names_from = type,
    values_from = sum_volume
  ) %>% 
  plot_ly(x = ~year, y = ~conventional, type = 'bar', name = 'Conventional') %>% 
  add_trace(y = ~organic, name = 'Organic') %>% 
  layout(yaxis = list(title = 'Volume Consumption'), barmode = 'stack', 
         title = "Conventional vs. Organic")

avo_df %>% 
  group_by(type) %>% 
  summarise(sum_volume = sum(total_volume)) %>% 
  plot_ly(labels = ~type, values = ~sum_volume, type = 'pie') %>% 
  layout(title = 'United States Avocado Consumption (2015-2018)')

```


```{r include=FALSE}
avo_df %>% 
  filter(year == 2015) %>% 
  group_by(region) %>% 
  summarise(region_volume = sum(total_volume)) %>% 
  mutate(
    code = state.abb[match(region,state.name)] 
  ) 

# the region we have is not state mostly ..
```

Time vs. Avocado Consumption by Region
```{r}
time_fig = 
  avo_df %>% 
  group_by(year, month, region) %>% 
  summarise(sum_volume = sum(total_volume)) %>% 
  ggplot(aes(x = month, y = sum_volume, color = region)) +
  geom_point(size = 0.5) +
  geom_smooth(size = 0.3, se = FALSE) + 
  facet_grid(~year) +
  labs(
    title = "Time vs. Avocado Consumption by Region",
    x = "Month",
    y = "Volume Consumption",
    color = "Region",
    caption = "https://hassavocadoboard.com/"
    ) + 
  scale_x_continuous(
    breaks = c(seq(1, 12, by = 3)),
    labels = c(seq(1, 12, by = 3))
    ) 

ggplotly(time_fig)
```

Time vs. Avocado Price by Region
```{r}
time_fig = 
  avo_df %>% 
  group_by(year, month, region) %>% 
  summarise(mean_price = mean(average_price)) %>% 
  ggplot(aes(x = month, y = mean_price, color = region)) +
  geom_point(size = 0.5) +
  geom_smooth(size = 0.3, se = FALSE) + 
  facet_grid(~year) +
  labs(
    title = "Time vs. Avocado Price by Region",
    x = "Month",
    y = "Average Price",
    color = "Region",
    caption = "https://hassavocadoboard.com/"
    ) + 
  scale_x_continuous(
    breaks = c(seq(1, 12, by = 3)),
    labels = c(seq(1, 12, by = 3))
    ) 

ggplotly(time_fig)
```



